Learning Excursion Sets of Vector-valued Gaussian Random Fields for Autonomous Ocean Sampling

### 2021-01-07 ans_to_reviewers/ans-to-2nd-rev.tex
Updates from Overleaf

\reply: We agree and have addressed this point in \aecom \textbf{1 - Reply to AE comment 4}.

@@

\vspace{5mm}
\reply:

@@

\vspace{5mm}
\reply:

@@

\vspace{5mm}
\reply: Line type will be tricky for Fig. 8, when lines are very close together especially in Fig.8b, and Fig.8c.

@@

\vspace{5mm}
\reply:

@@

\vspace{5mm}
\noindent \aecom \textbf{1 - Reply to AE comment 4}: 
I hank you for explaining the sampling limitations to me in your reply.  However, I think these limitations should also be noted in the narrative, and the sub-optimality of the resulting trajectory should be acknowledged.

@@

\vspace{5mm}
\reply:

@@

\vspace{5mm}
\reply:

>>>>

\reply: We agree and have addressed this point in \aecom \textbf{1}.

@@

\vspace{5mm}
\reply: We have elaborated on the specification of nugget in the revised version. We have addressed this point in \aecom \textbf{5}.

@@

\vspace{5mm}
\reply: We agree that the remarks of Ehrenberg are very useful! (It should be curriculum for graduate students.) Regarding the structure, we seriously considered your suggestions about the background in supplementary material, but in the end thought it was too much of a reduction to go through with. It would be very difficult to grasp the applied contribution of our paper if this background part was in the supplemental material, and only some readers take the extra effort of reading supplemental material. Based on the report of the reviewer, we did however go through the background and cut a few unnecessary and lengthy formulations in the revised version.

@@

\vspace{5mm}
\reply: We agree on all our points here, except Fig.8. Line type will be tricky for Fig. 8, when lines are very close together especially in Fig.8b, and Fig.8c. So we ask to consider that this figure is kept in color. Fig. 4, 5, and 6 have been made with a black-and-white compatible color-map, so they can be converted to black-and-white.

@@

\vspace{5mm}
\reply: We agree that this effort should be noted. We have added the following statement to the acknowledgements.

\begin{quote}
"We thank the anonymous reviewers for their careful reading of our manuscript and their many insightful comments and suggestions.``
\end{quote}

@@

\vspace{5mm}
\noindent \aecom \textbf{1 - Reply to AE comment 4}: 
I thank you for explaining the sampling limitations to me in your reply.  However, I think these limitations should also be noted in the narrative, and the sub-optimality of the resulting trajectory should be acknowledged.

@@

\vspace{5mm}
\reply: We have elaborated on the specification of nugget in the revised version. The sensor-provided standard deviations of the measurement tools are in the fourth decimal points, and in related data-sets we have not seen any nugget effect when measuring temperature and salinity in water samples. We believe that one challenge with the initial sampling line is that it was rather short and that it did no samples at very short distance. We made an empirical variogram from the actual survey data, which are shown below in Figure \ref{empVar}, and these curves support our statements of a smaller nugget.
\begin{figure}[!b] 
\centering
\includegraphics[width=0.99\textwidth]{Figures/field-trials/empVAR.png}
\caption{Empirial variogram from the two runs. }
\label{empVar}
\end{figure}

One reason to keep a significant nugget effect here is to account for the dynamic water masses that are not captured in our model, but we expect this to be very small in our limited survey time. Then again, between the two surveys, there is a change in the survey 1 and 2 paths caused by slightly different measurements at the two times of surveying. Another reason for including a nugget is the uncertainty in the AUV location which means it samples might not be allocated to its true position.

We further checked the sensitivity of the nugget in the sequential algorithm. This was done by running synthetic cases of sampling paths and data up to a certain stage, but then updating the model with different nugget effects in the Gaussian random field model. Figure X shows sampling paths and the radar plots for the AUV direction choice at one steps (top vs bottom) for two different nuggets. Straighter lines when the nugget is large.

In the revised manuscript we wrote another sentence on Fig 11: 'xx'

@@

\vspace{5mm}
\reply: We have improved our

### 2021-01-07 main_web.tex
Updates from Overleaf

Surprisingly less attention has been dedicated to sequential
strategies in the case of vector-valued observations. It has been long
acknowledged that co-Kriging could be updated efficiently in the context of sequential data assimilation \citep{Vargas-Guzman1999}, but sequential
strategies for estimating features of vector-valued random fields are
still in their infancy. \cite{LeGratiet.etal2015} used co-Kriging
based sequential designs to multi-fidelity computer codes and
\cite{Poloczek2017} used related ideas for multi-information source
optimization, but not for ES's like we do here. More relevant to our
setting, the PhD thesis \citep[][p.82]{stroh} mentions general
possibilities of stepwise uncertainty reduction strategies for ES's in
the context of designing fire simulations, yet outputs are mainly
assumed independent.

@@

The statistical model parameters were specified based on a short
preliminary survey where the AUV made an initial transect to determine
the trends in environmental conditions and correlation
structures. Based on the initial runs we get a reasonable idea of the temperature and salinity of river and ocean waters, and also specify the trend by linear regression, where both temperature and salinity were assumed to increase linearly with the west coordinate. Next, the residuals from the regression analysis were analyzed to specify the covariance parameters of the GRF model.
\begin{figure}[!b]
  \centering
  \subfigure[Residual plot.]{\includegraphics[width = 0.32\textwidth]{Figures/web/res_diag_a.pdf}\label{fig:parest_a}}
  \hfill
  \subfigure[Empirical CDF.]{\includegraphics[width = 0.32\textwidth]{Figures/web/res_diag_b.pdf}\label{fig:parest_b}}
  \hfill
  \subfigure[Empirical variogram.]{\includegraphics[width = 0.32\textwidth]{Figures/web/res_diag_c.pdf}\label{fig:parest_c}}
  \caption{Data analysis from a preliminary trial experiment using the
    AUV. \ref{fig:parest_a} Residual plot of temperature and salinity
    along with Gaussian contours. \ref{fig:parest_b} Empirical CDF
    (solid) of the quadratic form of the residuals along with the
    theoretical CDF (dashed) of the $\chi^2$ distribution with two
    degrees of freedom. \ref{fig:parest_c} Empirical variogram of the
    salinity and temperature data.}
\label{fig:parest}
\end{figure}
Fig.~\ref{fig:parest} summarizes diagnostic plots of this
analysis. Fig.~\ref{fig:parest_a} shows a cross-plot of temperature
and salinity residuals after the regression mean values of salinity and
temperature are subtracted from the data. This scatter-plot of joint
residuals indicates larger variability in salinity than temperature,
and a positive correlation ($0.5$) between the two variables. Based on
the fitted bivariate covariance model (ellipses in
Fig.~\ref{fig:parest_a}), we can compute the scalar quadratic form of
the residuals, and if the model is adequate they should be approximately $\chi^2_2$ distributed. Fig.~\ref{fig:parest_b} shows
the empirical CDF of the quadratic forms (solid) together with the
theoretical CDF of the $\chi^2_2$ distribution (dashed). The modeled and
theoretical curves are rather similar, which indicates that the Gaussian
model with constant spatial variance and correlation fits reasonably well. Fig.~\ref{fig:parest_c} shows the
empirical variogram of the residuals for temperature and salinity. The decay is similar for the two, and seems to be negligible
after about $150$ m.  The working assumption of a separable covariance
function is hence not unreasonable.

Based on the analysis in Fig.~\ref{fig:parest}, the resulting
parameters are given in Table \ref{tab:experiment_param}. The
regression parameters shown here are scaled to represent the east and
west boundaries of the domain as seen in the preliminary transect
data, and the thresholds are intermediate values. These parameter
values were then used in field trials where we explored the
algorithm's ability to characterize the river plume front separating
the river and fjord water masses.

@@

A Light AUV \citep{sousa2012lauv} (Fig.~\ref{fig:lauv}) equipped with
a 16 Hz Seabird Fastcat-49 conductivity, temperature, and depth (CTD)
sensor was used to provide salinity and temperature measurements.  The
AUV is a powered untethered platform that operates at $1$-$3$ m/s in
the upper water column. It has a multicore GPU NVIDIA Jetson TX1
(quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4
GB of 1600 MHz DRAM) for computation onboard.  The sampling algorithm
was built on top of the autonomous Teleo-Reactive EXecutive
(\textit{T-REX}) framework \citep{py10,Rajan12,Rajan12b}. We assume
that the measurements are conditionally independent because the
salinity is extracted from the conductivity sensor which is different
from the temperature sensor. We specify variance $0.25^2$ for both
errors, which is based on a middle ground between the nugget effect in
the empirical variogram and the sensor specifications.

@@

@@

The case study considers the upper water column in the river plume,
represented by a two dimensional grid. Extensions to three-dimensional
domains are not methodologically different, but likely approximate
calculations by concentrating numerical integration on terms in the
vicinity of the autonomous vehicle \citep{fossum18b}. While we did not
consider any temporal effects, which would be relevant on a larger
time scale, we do consider the extension to spatio-temporal modeling
and envision that advection-diffusion equations could be useful
\citep{sigrist2015stochastic,richardson2017sparsity}. For more complex
oceanographic phenomena, the methods will need to be extended to
non-Gaussian phenomena, possibly feature-based mixtures of Gaussian
processes which could potentially be run onboard augmented by
dynamical models. Running numerical models onboard a robotic vehicle
is currently infeasible, but high-resolution ocean models or remote
sensing data can be used to fit a more complex statistical model
\cite{davidson19}.

@@

TOF acknowledges support from the Centre for Autonomous Marine
Operations and Systems (AMOS), Center of Excellence, project number
223254, and the Applied Underwater Robotics Labortatory (AURLab). CT
and DG acknowledge support from the Swiss National Science Foundation,
project number 178858. JE and KR acknowledge support from Norwegian
research council (RCN), project number 305445. DG would like to
aknowledge support of Idiap Research Institute, his primary
affiliation in an early version of this manuscript. The authors would
also like to thank Niklas Linde of the University of Lausanne for
providing constructive feedback about this work, and members of the
NTNU AURLab for help with AUV deployments

>>>>

Surprisingly less attention has been dedicated to sequential
strategies in the case of vector-valued observations. It has been long
acknowledged that co-Kriging could be updated efficiently in the context of sequential data assimilation \citep{Vargas-Guzman1999}, but sequential
strategies for estimating features of vector-valued random fields are
still in their infancy. \cite{LeGratiet.etal2015} used co-Kriging
based sequential designs to multi-fidelity computer codes and
\cite{Poloczek2017} used related ideas for multi-information source
optimization, but not for ES's like we do here. More relevant to our
setting, \citet[p.82]{stroh} mentions general mentions general
possibilities of step-wise uncertainty reduction strategies for ES's in
the context of designing fire simulations, yet outputs are mainly
assumed independent.

@@

The statistical model parameters were specified based on experience as well as data gathered in a short
preliminary survey where the AUV made an initial transect to determine
the trends in environmental conditions, variance and correlation
structures. This gives a reasonable idea of the temperature and salinity of river and ocean waters. Both temperature and salinity were assumed to increase linearly with the west coordinate. Next, the residuals from the regression analysis were analyzed to specify the covariance parameters of the GRF model.
The resulting
parameters are given in Table \ref{tab:experiment_param}. The
regression parameters shown here are scaled to represent the east and
west boundaries of the domain as seen in the preliminary transect
data, and the thresholds are intermediate values. These parameter
values were then used in field trials where we explored the
algorithm's ability to characterize the river plume front separating
the river and fjord water masses.

@@

A Light AUV \citep{sousa2012lauv} (Fig.~\ref{fig:lauv}) equipped with
a 16 Hz Seabird Fastcat-49 conductivity, temperature, and depth (CTD)
sensor was used to provide salinity and temperature measurements.  The
AUV is a powered untethered platform that operates at $1$-$3$ m/s in
the upper water column. It has a multicore GPU NVIDIA Jetson TX1
(quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4
GB of 1600 MHz DRAM) for computation onboard.  The sampling algorithm
was built on top of the autonomous Teleo-Reactive EXecutive
(\textit{T-REX}) framework \citep{py10,Rajan12,Rajan12b}. We assume
that the measurements are conditionally independent because the
salinity is extracted from the conductivity sensor which is different
from the temperature sensor. We specify variance $0.25^2$ for both
errors. This is based on prior experience indicating that there can be some noise in temperature and salinity readings mostly because of dynamic water masses and uncertainty in AUV position. The initial survey transect was not sampled very densely so a reliable extraction of a nugget parameter from those data was not possible.

@@

\begin{figure}[!b]
  \centering
  \subfigure[Residual plot.]{\includegraphics[width = 0.32\textwidth]{Figures/web/res_diag_aN.eps}\label{fig:parest_a}}
  \hfill
  \subfigure[Empirical CDF.]{\includegraphics[width = 0.32\textwidth]{Figures/web/res_diag_bN.eps}\label{fig:parest_b}}
  \hfill
  \subfigure[Empirical variogram.]{\includegraphics[width = 0.32\textwidth]{Figures/web/res_diag_cN.eps}\label{fig:parest_c}}
  \caption{Data analysis from from AUV experiments. \ref{fig:parest_a} Residual plot of temperature and salinity
    along with specified Gaussian contours. \ref{fig:parest_b} Empirical CDF
    (solid) of the quadratic form of the residuals along with the
    modeled theoretical CDF (dashed) of the $\chi^2$ distribution with two
    degrees of freedom. \ref{fig:parest_c} Scaled empirical variograms of the
    salinity and temperature data.}
\label{fig:parest}
\end{figure}

Of course, the initial 
Fig.~\ref{fig:parest} summarizes diagnostic plots of this
analysis. Fig.~\ref{fig:parest_a} shows a cross-plot of temperature
and salinity residuals after the regression mean values of salinity and
temperature are subtracted from the data. This scatter-plot of joint
residuals indicates that the correlation between the two variables might be slightly larger than the initially specified $0.5$ and that uncertainty in temperature was set a bit too high. Based on
the fitted bivariate covariance model (ellipses in
Fig.~\ref{fig:parest_a}), we can compute the scalar quadratic form of
the residuals, and if the model is adequate they should be approximately $\chi^2_2$ distributed. Fig.~\ref{fig:parest_b} shows
the empirical CDF of the quadratic forms (solid) together with the
theoretical CDF of the $\chi^2_2$ distribution (dashed). The empirical and
theoretical curves are not very different, which indicates that the specified Gaussian
model fits reasonably well. Fig.~\ref{fig:parest_c} shows the
empirical variogram of the scaled residuals for temperature and salinity. The decay is similar for the two, and seems to be negligible
after about $150$ m as was specified. The working assumption of a separable covariance
function is hence not unreasonable.

Overall, the spatial Gaussian model representation is of course a simplification of the changing water conditions, and even within the model formulation, the parameter specification could likely have been improved. 
However, we do not expect minor changes to have caused much difference in sampling paths.

@@

The case study considers the upper water column in the river plume,
represented by a two dimensional grid. Extensions to three-dimensional
domains are not methodologically different, but likely approximate
calculations by concentrating numerical integration on terms in the
vicinity of the autonomous vehicle \citep{fossum18b}. While we did not
consider any temporal effects, which would be relevant on a larger
time scale, we do consider the extension to spatio-temporal modeling
and envision that advection-diffusion equations could be useful
\citep{sigrist2015stochastic,richardson2017sparsity}. The sampling capabilities are restricted by the sensors onboard the AUV, the computational load and navigation resources. The sampling of temperature and salinity variables can be much higher than the grid resolution, but, due to computational and navigation resources
available on the AUV, assimilation into a discretized fixed grid is necessary.For more complex
oceanographic phenomena, the methods will need to be extended to
non-Gaussian phenomena, possibly feature-based mixtures of Gaussian
processes which could potentially be run onboard augmented by
dynamical models. Running numerical models onboard a robotic vehicle
is currently infeasible, but high-resolution ocean models or remote
sensing data can be used to fit a more complex statistical model
\cite{davidson19}.

@@

TOF acknowledges support from the Centre for Autonomous Marine
Operations and Systems (AMOS), Center of Excellence, project number
223254, Nansen Legacy Program, project number 276730, and the Applied Underwater Robotics Labortatory (AURLab). CT
and DG acknowledge support from the Swiss National Science Foundation,
project number 178858. JE and KR acknowledge support from Norwegian
research council (RCN), project number 305445. DG would like to
aknowledge support of Idiap Research Institute, his primary
affiliation in an early version of this manuscript. We thank the anonymous    reviewers for their careful reading of our manuscript and their many insightful comments and suggestions. The authors would also like to thank Niklas Linde of the University of Lausanne for providing constructive feedback about this work, and members of the NTNU AURLab for help with AUV deployments
